Homework

### 2016-07-04 Macro Impacts.tex
Additional short discussion

\pagebreak
\section{Our Argument}
As a Macro passes through the earth, there are several phases to its journey. The first is its passage through the atmosphere. The second is its crossing the boundary from air to earth, which we suspect will be relatively uneventful, since the air is 1000$\times$ lower density than earth, and thus the air pressure would displace relatively little of earth. The third is the transit of the macro through the earth. The shockwave it produces should be generated primarily by elastic and inelastic scattering of baryonic matter off the Macro. If we estimate that the macro produces an initial overpressure in a tube of its own radius, we can calculate the initial overpressure from the total energy deposition and its cross section. Given that the macro is dense enough, its loss of momentum will be approximately linear, and we can approximate its energy loss as being approximately constant. Estimates have already been made of the stopping power, so the total energy deposition will just be
\begin{align}
\abs{\td{E}{x}{}}h=E_{\te{total}}=\f{\lambda+2\mu}{K^2}4\pi^2p_0^2\sigma_Xh
\end{align}
whence the initial overpressure is
\begin{align}
p_0=\f{K}{2\pi}\sqrt{\abs{\td{E}{x}{}}\f{1}{\sigma_X(\lambda+2\mu)}}
\end{align}
It is not appropriate to use the linear regime until the overpressure is close to 0 compared to some pressure, which we suspect may be $p_{\te{non-linear}}\approx60\te{GPa}$. It is reasonable to assume that the elastic limit holds when the pressures are an order of magnitude smaller than $p_{\te{non-linear}}$, denoted $p_{\te{linear}}\approx6\te{GPa}$. (This stuff needs to be checked). Note that energy losses due to heating, i.e. anelastic frequency attenuation, will dilute the energy density of the shock. However, for a conservative  estimate, we only consider the geometric attenuation, which will dilute the energy density as $r^{-1}$, where $r$ is the radial distance from the Macro's trajectory to the shock front. The shape of the shockwave will rapidly achieve a right-triangular profile. Geometric considerations show that the overpressure will go as $r^{-1}$. Thus, suppose that the peak pressure in the shock is $\bar p$. The overpressure is then given by
\begin{align}
p=\bar p\f{r-r_0}{\Delta r}\ps{\theta(r_0+\Delta r-r)-\theta(r+\Delta r)}\ps{\theta(z+h/2)-\theta(z-h/2)}
\end{align}
where $\Delta r$ is the radial length of the shock and $r_0$ is its base. In the limit where $h$ is much larger than all other fixed lengths, we have (after a large but easy mathematica calculation (triangle pressure.nb))
\begin{align}
E_{\te{propagated}}&\approx \f{\lambda+2\mu}{K^2}\f{h \bar p^2\pi^2r_0^6}{8\Delta r^2\lambda_0^2}\,.
\end{align}
Generously, we assume that the energy of the wave stayed constant through its non-linear evolution, and calculate the fraction energy deposited into the low frequency modes
\begin{align}
\Xi&=\f{h \bar p^2\pi^2r_0^6}{8\Delta r^2\lambda_0^2}\f{1}{4\pi^2p_0^2\sigma_Xh}\,,\\
&=\f{1}{32}\f{\bar p^2}{p_0^2}\f{r_0^6}{\Delta r^2\sigma_X\lambda_0^2}\,,\\
&=\f{\pi^2(\lambda+2\mu)}{8K^2}\f{\bar p^2}{\abs{\td{E}{x}{}}}\f{r_0^6}{\Delta r^2\lambda_0^2}\,.
\end{align}
$\bar p$ can be taken to be the maximum stress such that the elastic limit holds i.e. the elastic limit. Some reading suggests that this is about 13 GPa. $K$ and $\lambda+2\mu$ can be taken to be on the order of $10^{10}$ GPa (obviously we should look into this). $\Delta r$ should be very small, on the order of the radius of the macro. $\lambda_0$ is about $6$ to $10$ kilometers, and is the wavelength at which modes propagate without significant attenuation. $r_0$ can be determined by geometric means. A first guess yields
\begin{align}
\Xi&\approx\f{K^4}{(\lambda+2\mu)^2}\abs{\td{E}{x}{}}^2\f{1}{512\bar p^4\lambda_0^2\sigma_X}\,,\\
&\approx10^{-18}
\end{align}
This is within one order of magnitude of the answer gotten from (24).

>>>>

\section{Our Argument}
As a Macro passes through the earth, there are several phases to its journey. The first is its passage through the atmosphere. The second is its crossing the boundary from air to earth, which we suspect will be relatively uneventful, since the air is 1000$\times$ lower density than earth, and thus the air pressure would displace relatively little of earth. The third is the transit of the macro through the earth. The shockwave it produces should be generated primarily by elastic and inelastic scattering of baryonic matter off the Macro. If we estimate that the macro produces an initial overpressure in a tube of its own radius, we can calculate the initial overpressure from the total energy deposition and its cross section. Given that the macro is dense enough, its loss of momentum will be approximately linear, and we can approximate its energy loss as being approximately constant. Estimates have already been made of the stopping power, so the total energy deposition will just be
\abs{\td{E}{x}{}}h=E_{\te{total}}=\f{\lambda+2\mu}{K^2}4\pi^2p_0^2\sigma_Xh
whence the initial overpressure is
p_0=\f{K}{2\pi}\sqrt{\abs{\td{E}{x}{}}\f{1}{\sigma_X(\lambda+2\mu)}}
It is not appropriate to use the linear regime until the overpressure is close to 0 compared to some pressure, which we suspect may be $p_{\te{non-linear}}\approx60\te{GPa}$. It is reasonable to assume that the elastic limit holds when the pressures are an order of magnitude smaller than $p_{\te{non-linear}}$, denoted $p_{\te{linear}}\approx6\te{GPa}$. (This stuff needs to be checked). Note that energy losses due to heating, i.e. anelastic frequency attenuation, will dilute the energy density of the shock. However, for a conservative  estimate, we only consider the geometric attenuation, which will dilute the energy density as $r^{-1}$, where $r$ is the radial distance from the Macro's trajectory to the shock front. The shape of the shockwave will rapidly achieve a right-triangular profile. Geometric considerations show that the overpressure will go as $r^{-1}$. Thus, suppose that the peak pressure in the shock is $\bar p$. The overpressure is then given by
p=\bar p\f{r-r_0}{\Delta r}\ps{\theta(r_0+\Delta r-r)-\theta(r+\Delta r)}\ps{\theta(z+h/2)-\theta(z-h/2)}
where $\Delta r$ is the radial length of the shock and $r_0$ is its base. In the limit where $h$ is much larger than all other fixed lengths, we have (after a large but easy mathematica calculation (triangle pressure.nb))
E_{\te{propagated}}&\approx \f{\lambda+2\mu}{K^2}\f{h \bar p^2\pi^2r_0^6}{8\Delta r^2\lambda_0^2}\,.
Generously, we assume that the energy of the wave stayed constant through its non-linear evolution, and calculate the fraction energy deposited into the low frequency modes
\Xi&=\f{h \bar p^2\pi^2r_0^6}{8\Delta r^2\lambda_0^2}\f{1}{4\pi^2p_0^2\sigma_Xh}\,,\\
&=\f{1}{32}\f{\bar p^2}{p_0^2}\f{r_0^6}{\Delta r^2\sigma_X\lambda_0^2}\,,\\
&=\f{\pi^2(\lambda+2\mu)}{8K^2}\f{\bar p^2}{\abs{\td{E}{x}{}}}\f{r_0^6}{\Delta r^2\lambda_0^2}\,.
$\bar p$ can be taken to be the maximum stress such that the elastic limit holds i.e. the elastic limit. Some reading suggests that this is about 13 GPa. $K$ and $\lambda+2\mu$ can be taken to be on the order of $10^{10}$ GPa (obviously we should look into this). $\Delta r$ should be very small, on the order of the radius of the macro. $\lambda_0$ is about $6$ to $10$ kilometers, and is the wavelength at which modes propagate without significant attenuation. $r_0$ can be determined by geometric means. A first guess yields
\Xi&\approx\f{K^4}{(\lambda+2\mu)^2}\abs{\td{E}{x}{}}^2\f{1}{512\bar p^4\lambda_0^2\sigma_X}\,,\\
&\approx10^{-18}
This is within one order of magnitude of the answer gotten from (24). Note, my estimate for $r_0$ turned out to be about 1 kilometer. I'm still not entirely sure of the correct way to calculate $r_0$.
